

1. Load libraries:

```{r, message = FALSE}
require(biomaRt)
require(ggplot2)
require(dplyr)
require(lintr)
require(eulerr)

2. Load ensembl databases:

```{r, message = FALSE}
listMarts()

ensembl <- useMart("ensembl")

datasets <- listDatasets(ensembl)

ensembl <- useDataset("strutta_gene_ensembl",
                     mart = ensembl)
```
3. Extract NCBI symbols:

```{r, message = FALSE}
## Identify NCBI gene ids:
salmo_ncbi <- getBM(attributes = c('ensembl_gene_id',
                                   'entrezgene_id',
                                   'chromosome_name',
                                   'start_position',
                                   'end_position'),
                    mart = ensembl)
```

4. Subset InterProScan domains assigned to brown trout immune ohnologs:

```{r, message = FALSE}
salmo_interpro <- getBM(attributes = c('ensembl_gene_id',
                                   'chromosome_name',
                                   'start_position',
                                   'end_position',
                                   'interpro',
                                   'interpro_description'),
                    mart = ensembl)

salmo_interpro_df <- as.data.frame(table(salmo_interpro$ensembl_gene_id))
```

5. Read in list of immune ohnologs:

```{r, message = FALSE}
immune_ohnologs <- read.table(file = "input/ancestral_copy_plus_protein_length.txt",
                              header = FALSE)
immune_ohnologs$V5 <- gsub(pattern = "[.]1",
                           replacement = "",
                           immune_ohnologs$V5)
immune_ohnologs$V4 <- gsub(pattern = "[.]1",
                           replacement = "",
                           immune_ohnologs$V4)

immune_ohnologs$con_domains <- salmo_interpro_df[match(immune_ohnologs$V5,
                        salmo_interpro_df$Var1), ]$Freq
immune_ohnologs$div_domains <- salmo_interpro_df[match(immune_ohnologs$V4,
                        salmo_interpro_df$Var1), ]$Freq

## Compare mean number of functional domains:
wilcox.test(immune_ohnologs$con_domains,
            immune_ohnologs$div_domains)

t.test(immune_ohnologs$V7,
            immune_ohnologs$V8)

t.test(immune_ohnologs$V1,
            immune_ohnologs$V2)

ggplot(data = immune_ohnologs,
       aes(x = V7,
           y = div_domains,
           colour = "blue")) +
        geom_point() +
        geom_point(data = immune_ohnologs,
                   aes(x = V8,
                       y = con_domains,
                       colour = "red"))

immune_ohnologs_domains_diff <- subset(immune_ohnologs,
          con_domains != div_domains)

nrow(subset(immune_ohnologs_domains_diff,
       con_domains < div_domains))
```
6. Read in clusters and check overlap.

```{r, message = FALSE}
cluster_1 <- scan(file = "data/Clusters_Objects_cluster_1_gene_list.txt",
                  as.character())
cluster_1_df <- as.data.frame(cbind(locus = cluster_1,
                      cluster = "1"))
cluster_2 <- scan(file = "data/Clusters_Objects_cluster_2_gene_list.txt",
                  as.character())
cluster_2_df <- as.data.frame(cbind(locus = cluster_2,
                      cluster = "2"))
cluster_3 <- scan(file = "data/Clusters_Objects_cluster_3_gene_list.txt",
                  as.character())
cluster_3_df <- as.data.frame(cbind(locus = cluster_3,
                      cluster = "3"))
cluster_4 <- scan(file = "data/Clusters_Objects_cluster_4_gene_list.txt",
                  as.character())
cluster_4_df <- as.data.frame(cbind(locus = cluster_4,
                      cluster = "4"))
cluster_5 <- scan(file = "data/Clusters_Objects_cluster_5_gene_list.txt",
                  as.character())
cluster_5_df <- as.data.frame(cbind(locus = cluster_5,
                      cluster = "5"))
cluster_6 <- scan(file = "data/Clusters_Objects_cluster_6_gene_list.txt",
                  as.character())
cluster_6_df <- as.data.frame(cbind(locus = cluster_6,
                      cluster = "6"))
cluster_7 <- scan(file = "data/Clusters_Objects_cluster_7_gene_list.txt",
                  as.character())
cluster_7_df <- as.data.frame(cbind(locus = cluster_7,
                      cluster = "7"))
cluster_8 <- scan(file = "data/Clusters_Objects_cluster_8_gene_list.txt",
                  as.character())
cluster_8_df <- as.data.frame(cbind(locus = cluster_8,
                      cluster = "8"))
cluster_9 <- scan(file = "data/Clusters_Objects_cluster_9_gene_list.txt",
                  as.character())
cluster_9_df <- as.data.frame(cbind(locus = cluster_9,
                      cluster = "9"))

## Generate a combined plot:
combined_df <- rbind(cluster_1_df,
                     cluster_2_df,
                     cluster_3_df,
                     cluster_4_df,
                     cluster_5_df,
                     cluster_6_df,
                     cluster_7_df,
                     cluster_8_df,
                     cluster_9_df)

combined_df$locus <- gsub(pattern = "[.].*",
                          replacement = "",
                          combined_df$locus)

head(na.omit(combined_df[match(combined_df$locus,
                  immune_ohnologs_domains_diff$V5), ]))
```

6. Run lintr:

```{r, message = FALSE}
lintr::lint(file = "identify_immune_genes_ensembl_ssalar.Rmd")
```
