--- 
title: "Salmonid immune study"
output: plot_go_term_barchart.html
author: Joe Colgan (joscolgan)
---

## Introduction:  
This script is for visualisation of the output of gene ontology (GO) analyses using the R package TopGo.
This script creates a barchart of the number of significant genes annotated to a particular GO term, with an associated p-value. This script takes an input of .csv files containing an output table of GO analyses.

```{r, message = FALSE}
# Load libraries; install from scratch if needed
libraries <- c("ggplot2",
               "lattice",
               "ggpubr")
for (lib in libraries) {
    if (require(package = lib, character.only = TRUE)) {
        print("Successful")
    } else {
        print("Installing")
        source("https://bioconductor.org/biocLite.R")
        library(lib, character.only = TRUE)
    }
}
## Create output directory:
dir.create("results")
```

## Step One: Load input files

```{r, message = FALSE}
## Read in input:
results <- "./results/strutta_drerio_immmune/input_for_barchart/"
## Read in biological process terms:
sig_bp_terms <- read.table(file = paste(results,
                                    "sig_BP_terms_1_to_more_than_2_input_for_plot.txt",
                                    sep = ""),
                                    header = FALSE,
                           col.names =  c("go_term",
                                        "go_description",
                                        "go_number",
                                        "adjusted_p"))
                       
## Subset significant values:
sig_bp_terms <- subset(sig_BP_terms,
                   adjusted_p < 0.05)
sig_bp_terms$category <- "1:1"

sig_bp_terms$amended_terms <- paste(sig_bp_terms$go_description,
                                   "_(",
                                   sig_bp_terms$go_number,
                                   ")",
                                   sep = "")
sig_bp_terms$amended_terms <- gsub(pattern = "_",
                                  replacement = " ",
                                  sig_bp_terms$amended_terms)
## Subset genes with adjust p value less than 0.01
sig_bp_terms <- subset(x = sig_bp_terms,
                      adjusted_p < 0.01)
## Log transform p value:
sig_bp_terms$log_10 <- -log10(x = sig_bp_terms$adjusted_p)
```

Generate plot:

```{r, message = FALSE}
## Function for plotting barchart:
plot <- ggbarplot(sig_BP_terms,
                          x = "amended_terms",
                          y = "log_10",
                          position = position_dodge(0.1),
                          fill = "category",
                          color = NULL,
                          palette = "jco",
                          sort.val = "asc",
                          sort.by.groups = TRUE,
                          ylab = "-log10(p)",
                          xlab = "Gene Ontology term",
                          legend.title = "Gene Ontology",
                          lab.col = "black",
                          lab.size = 4,
                          lab.vjust = 0.5,
                          lab.hjust = 1,
                          legend = "top",
                          rotate = TRUE,
                          ggtheme = theme_minimal())

## Make font bigger and bold:
plot <- plot +
                scale_y_continuous(expand = c(0, 0)) +
                theme(axis.text = element_text(size = 10),
                      axis.title.x = element_text(size = 12,
                                                  face = "bold"),
                      axis.title.y = element_text(size = 12,
                                                  face = "bold"),
                      axis.text.y = element_text(size = 10,
                                                 face = "bold"),
                      axis.text.x = element_text(size = 10),
                      legend.position = "none") +
                expand_limits(y = 10) +
                geom_hline(yintercept = 1.301,
                           linetype = "dashed",
                           colour = "black")
## Update colours for plotting
plot <- plot +
        scale_fill_manual(values = c("light blue"))

## Save picture:
ggsave(file = "results/1_more_than_2_go_terms_barchart.pdf",
       dpi = 600,
       height = 20,
       width = 18)
```

Run lintr:

```{r, message =  FALSE}
lintr::lint(file = "plot_go_term_barchart.Rmd")
```
